\cleardoublepage
%\pagebreak
\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}
\chapter*{Acknowledgments}

We would like to give our sincerest thanks to all the people who have been involved in making this thesis work possible. Thanks to Anders Ynnerman for giving us this opportunity. Thanks to Carter Emmart for being an inspiring and driving force for the project and for sharing his passion in bringing knowledge and interest in astronomy to the general public. Thanks to Vivian Trakinski for making us feel needed and useful within the Openspace project and within the museum.

Thanks to Alexander Bock for his dedication in the project and the support he has given as a mentor along with Emil Axelsson during the whole project. Thanks to all the people in the OpenSpace team, including our peers Michael and Sebastian, which have been both inspiring, helpful and enjoyable to work and share the experience with.

We would like to thank all the people we have met during our time at AMNH. Kayla, Eozin, Natalia have not only been our trusted lunch mates but also great friends outside of work.
Thanks to Jay for all the hardware support, and also all the rest of the people at the Science Bulletins and Rose Center Engineering for being so welcoming and helpful.

We would also like to thank Lucian Plesea for his expert support in mapping services together with Vikram Singh for setting up the map servers we could use in our software. Also, thanks to Jason, Ally and David for providing us with high resolution Mars imagery data that we could use for rendering.

A big thank you to Masha and the rest of the CCMC team as well as Ryan, who made our visit to NASA Goddard Space Flight Center the best experience possible by inspiring us and giving us insight in parts of NASA's space science.

All our friends and family who travelled from Sweden to visit us in New York, we're happy for sharing a great time with you during our leisure. 

Last but not least we are very happy to have made new great friends outside of the thesis work during our stay in the United States. You have made this experience even more enjoyable.

\newpage
