\cleardoublepage
%\pagebreak
\phantomsection
\addcontentsline{toc}{chapter}{References}
\begin{thebibliography}{99}

%\bibitem{citation-1-name-here}<Name of the reference here>,\ \url{<url here>}

%@book{Cozzi11,
%    author = {Patrick Cozzi and Kevin Ring},
%    title = {3{D} {E}ngine {D}esign for {V}irtual {G}lobes},
%    month = {June},
%    year = {2011},
%    edition = {1st},
%    isbn = {978-1568817118},
%    publisher = {CRC Press}
%    note = {\url{http://www.virtualglobebook.com}}
%}

\bibitem{openspace}
  \emph{OpenSpace},
  \\
  \url{http://openspaceproject.com/},
  (accessed 2016-11-16)

\bibitem{cozzi11}
  Patrick Cozzi and Kevin Ring,
  \emph{3D Engine Design for Virtual Globes},
  CRC Press,
  1st edition,
  2011. \\
  \url{http://www.virtualglobebook.com}
  (accessed 2016-08-03)

\bibitem{clasen06}
  Malte Clasen and Hans-Christian Hege,
  \emph{Terrain rendering using spherical clipmaps},
  In Proceedings of the Eighth Joint Eurographics / IEEE VGTC conference on Visualization (EUROVIS'06),
  Aire-la-Ville, Switzerland,
  2006. \\
  \doi{ http://dx.doi.org/10.2312/VisSym/EuroVis06/091-098},
  (accessed 2016-11-16)

\bibitem{dimi15}
  Aleksandar M. Dimitrijevi\'{c} and Dejan D. Ran\v{c}i\'{c}
  \emph{Ellipsoidal Clipmaps - A planet-sized terrain rendering algorithm},
  In Proceedings of the Eighth Joint Eurographics / IEEE VGTC conference on Visualization (EUROVIS'06),
  Comput. Graph. 52, C (November 2015), 43-61,
  2015. \\
  \doi{ http://dx.doi.org/10.1016/j.cag.2015.06.006},
  (accessed 2016-11-01)

\bibitem{tanner98}
  Christopher C. Tanner, Christopher J. Migdal, and Michael T. Jones,
  \emph{The clipmap: a virtual mipmap},
  In Proceedings of the 25th annual conference on Computer graphics and interactive techniques (SIGGRAPH '98),
  New York, NY, USA, 151-158,
  1998. \\
  \doi{ http://dx.doi.org/10.1145/280814.280855},
  (accessed 2016-11-01)

\bibitem{losasso04}
  Frank Losasso and Hugues Hoppe
  \emph{Geometry Clipmaps: Terrain Rendering Using Nested Regular Grids},
  ACM Trans. Graphics (SIGGRAPH), 23(3),
  2004.
  
\bibitem{wms06}
  Open Geospatial Consortium Inc,
  \emph{OpenGIS Web Map Server Implementation Specification},
  2006.

\bibitem{wmts10}
  Open Geospatial Consortium Inc,
  \emph{OpenGIS Web Map Tile Service Implementation Standard},
  2010.

\bibitem{kemen12}
  Brano Kemen,
  \emph{Maximizing Depth Buffer Range and Precision},
  Article for the Outerra software,
  2012. \\
  \url{http://outerra.blogspot.se/2012/11/maximizing-depth-buffer-range-and.html}
  (accessed 2016-11-14)

\bibitem{outerraprocedural09}
  Brano Kemen,
  \emph{Procedural terrain algorithm visualization},
  2009, \\
  \url{http://outerra.blogspot.se/2009/02/procedural-terrain-algorithm.html}
  (accessed 2016-11-16)

\bibitem{terragen}
  Terragen - Planetside Software,
  \emph{Terragen Wiki - Procedural Data},
  2013, \\
  \url{http://planetside.co.uk/wiki/index.php?title=Procedural_Data},
  (accessed 2016-11-16)

\bibitem{roam}
  Mark Duchaineau, Murray Wolinsky, David E. Sigeti, Mark C. Miller, Charles Aldrich, and Mark B. Mineev-Weinstein.
  \emph{ROAMing Terrain: Real- Time Optimally Adapting Meshes},
  In Proceedings of the 8th Conference on Visualization \#97, pp. 81-88. Los Alamitos, CA: IEEE Computer Society,
  1997, \\
  \url{}

\bibitem{gdal}
  The Open Source Geospatial Foundation,
  \emph{GDAL - Geospatial Data Abstraction Library},
  GDAL Website,
  %2012.
  \\
  \url{http://www.gdal.org},
  (accessed 2016-11-16)

\bibitem{htm}
  Hierarchical Triangular Mesh,
  \emph{Sloan Digital Sky Survey},
  %,
  2007. \\
  \url{http://www.skyserver.org/HTM/}
  (accessed 2016-11-12)
  
\bibitem{toast}
  Jonathan Fay,
  \emph{WorldWide Telescope Projection Reference},
  %,
  %2012.
  \\
  \url{http://www.worldwidetelescope.org/docs/worldwidetelescopeprojectionreference.html},
  (accessed 2016-11-16)

\bibitem{spice}
  NASA's Navigation and Ancillary Information Facility,
  \emph{An Overview of SPICE, NASA's Ancillary Data System for Planetary Missions},
  %,
  2016. \\
  \url{http://naif.jpl.nasa.gov/pub/naif/toolkit_docs/Tutorials/pdf/individual_docs/03_spice_overview.pdf},
  (accessed 2016-11-16)

\bibitem{battersby14}
  Sarah E. Battersby, Michael P. Finn, E. Lynn Usery and, Kristina H. Yamamoto,
  \emph{Implications of Web Mercator and Its Use in Online Mapping},
  Cartographica 49:2, 2014, pp. 85-101,
  2014. \\
  \doi{ http://dx.doi.org/10.3138/carto.49.2.2313},
  (accessed 2016-11-11)

\bibitem{newton87}
  Isaac Newton,
  \emph{Philosophiae naturalis principia mathematica},
  J. Societatis Regiae ac Typis J. Streater,
  University of California,
  2011. \\
  \url{https://books.google.se/books?id=-dVKAQAAIAAJ},
  (accessed 2016-11-11)
  
\bibitem{epsg}
  European Petroleum Survey Group,
  %\emph{WorldWide Telescope Projection Reference},
  %,
  %2012.
  \\
  \url{http://www.epsg.org},
  (accessed 2016-11-11)
  
\bibitem{spatialref}
  Spatial reference list,
  %\emph{WorldWide Telescope Projection Reference},
  %,
  %2012.
  \\
  \url{http://spatialreference.org/ref/},
  (accessed 2016-08-10)


\bibitem{tms}
  Open Source Geospatial Foundation,
  \emph{Tile Map Service Specification},
  %,
  %2012.
  \\
  \url{https://wiki.osgeo.org/wiki/Tile_Map_Service_Specification},
  (accessed 2016-08-17)
  
  \bibitem{arcgis}
  ESRI,
  \emph{Arcgis Online},
  %,
  %2012.
  \\
  \url{https://www.arcgis.com},
  (accessed 2016-08-17)

  \bibitem{onmoon}
  NASA Jet Propulsion Laboratory,
  \emph{Onmoon Web Interface},
  California Institute of Technology
  %,
  %2012.
  \\
  \url{http://onmoon.jpl.nasa.gov},
  (accessed 2016-08-17)
  
  \bibitem{segal10}
  Mark Segal and Kurt Akeley,
  \emph{The OpenGL Graphics System: A Specification (Version 4.0 (Core Profile) - March 11, 2010)},
  CRC Press,
  1st edition,
  2010.
  
  \bibitem{sos}
  National Oceanic and Atmospheric Administration,
  \emph{Science On a Sphere},
  %,
  %2012.
  \\
  \url{http://sos.noaa.gov/},
  (accessed 2016-08-17)
  
  \bibitem{spaceengine}
  Vladimir Romanyuk,
  \emph{Space Engine Website},
  %,
  %2012.
  \\
  \url{http://en.spaceengine.org},
  (accessed 2016-08-17)
  
  \bibitem{outerra}
  Brano Kemen and Laco Hrabcak,
  \emph{Outerra Website},
  %,
  %2012.
  \\
  \url{http://www.outerra.com},
  (accessed 2016-08-17)
  
  \bibitem{googlemaps}
  \emph{Google Maps},
  %,
  %2012.
  \\
  \url{http://maps.google.se},
  (accessed 2016-08-17)
  
  \bibitem{uniview}
  SCISS AB,
  \emph{Uniview},
  %,
  %2012.
  \\
  \url{http://sciss.se/uniview},
  (accessed 2016-08-17)
  
  \bibitem{wwt}
  World Wide Telescope,
  \emph{American Astronomical Society},
  %,
  %2012.
  \\
  \url{http://www.worldwidetelescope.org},
  (accessed 2016-08-17)
  
  \bibitem{digistar}
  Evans \& Sutherland,
  \emph{Digistar},
  %,
  %2012.
  \\
  \url{http://www.es.com/Digistar},
  (accessed 2016-08-17)
    
  \bibitem{digitalsky}
  Sky-Skan,
  \emph{DigitalSky},
  %,
  %2012.
  \\
  \url{https://www.skyskan.com/products/ds},
  (accessed 2016-08-17)
    
  \bibitem{exoplanets}
  Sasha Hinkley,
  \emph{Directly Imaging Exoplanets},
  Astrophysics Group, University of Exeter,
  2016 \\
  \url{https://palereddot.org/directly-imaging-exoplanets/},
  (accessed 2016-08-18)
  
  \bibitem{ieee754}
  The Institute of Electrical and Electronics Engineers
  \emph{IEEE Standard for Binary Floating-Point Arithmetic},
  The Institute of Electrical and Electronics Engineers, Inc,
  New York,
  1985. \\
  \url{http://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=30711},
  (accessed 2016-11-18)
  
  \bibitem{ghoul}
  \emph{The General Helpful Open Utility Library},
  \\
  \url{https://github.com/OpenSpace/Ghoul},
  (accessed 2016-11-23)
  
  \bibitem{viking}
  U.S. Geological Survey,
  \emph{Mars Viking MDIM21 ClrMosaic global 232m},
  \\
  \url{http://astrogeology.usgs.gov/search/map/Mars/Viking/MDIM21/Mars_Viking_MDIM21_ClrMosaic_global_232m},
  (accessed 2016-11-23)
  
  \bibitem{ctx}
  \emph{Online Map Viewer for MRO CTX Patches},
  \\
  \url{http://viewer.mars.asu.edu/viewer/ctx#T=0},
  (accessed 2016-11-23)
  
  \bibitem{ctx}
  \emph{Online Map Viewer for MRO CTX Patches},
  \\
  \url{http://viewer.mars.asu.edu/viewer/ctx#T=0},
  (accessed 2016-11-23)
  
  \bibitem{hirise}
  Lunar \& Planetary Laboratory
  \emph{HiRISE Digital Terrain Models},
  \\
  \url{http://www.uahirise.org//dtm/},
  (accessed 2016-11-23)
  
  \bibitem{healpix}
  NASA, Jet Propulsion Laboratory,
  \emph{HEALPix},
  \\
  \url{http://healpix.jpl.nasa.gov/index.shtml},
  (accessed 2016-11-23)
  
  \bibitem{mapprojections}
  \emph{Map Projection Images},
  \\
  \url{https://en.wikipedia.org/wiki/List_of_map_projections},
  (accessed 2016-11-25)
  
  \bibitem{worldimagery}
  ESRI, ArcGIS Online,
  \emph{World Imagery},
  \\
  \url{https://www.arcgis.com/home/item.html?id=10df2279f9684e4a9f6a7f08febac2a9},
  (accessed 2016-11-25)
  
  \bibitem{imageryworld2d}
  ESRI, ArcGIS Online,
  \emph{ESRI Imagery World 2D WGS84},
  \\
  \url{https://www.arcgis.com/home/item.html?id=37b8a8b1014747deb5e15472d06d0da9},
  (accessed 2016-11-25)
  
\end{thebibliography}
